\section{Evaluation} \label{sec:evaluation}
This section presents the evaluation of the implemented Rocq proof checker and compares it with the baseline Imandra proof checker described in \cite{certified_proof_checker}. The evaluation is structured along two complementary dimensions: capability and maintainability.

\subsection{Capability Evaluation}

The Rocq implementation achieves functional parity with the Imandra baseline in terms of core features: both support ReLU activation constraints, correctly parse Marabou proof certificates and formalise their representations in the target language, include a complete proof checker, and are accompanied by a formal soundness theorem.

Nevertheless, the Rocq implementation is missing some functionalities. First, the Imandra proof checker supports the use of theory lemmas for dynamic bound tightening, which are fundamental for scaling to larger verification problems, although their soundness is not proven. Finally, the most important missing feature is the ability to execute the proof checker. The reason is the proof checker's high reliance on inefficient data types such as MathComp matrices \cite{refinement_algebra_rocq}. As a result, executing the checker becomes infeasible due to excessive memory consumption.

Despite the drawback that the proof checker cannot currently be executed and therefore cannot be used in real-world scenarios, we have identified multiple plausible solutions in \autoref{sec:future-work}. These solutions should help prevent the proof checker from incurring unreasonable memory usage during execution and enable it to correctly establish the soundness of a proof certificate.

\subsection{Maintainability Evaluation}
The maintainability evaluation focuses mainly on the codebases of both soundness implementations. Since both Gallina and Imandra (a subset of OCaml) share very similar syntax, we therefore skip the comparison between code related to both proof checkers. The maintainability of the code is assessed through both quantitative and qualitative analyses. In the quantitative analysis, we compare and analyse the overall formalisation complexity through metrics such as lines of code (L.O.C.), auxiliary lemmas, and dependencies. In the qualitative analysis, we compare the overall proof flow and provide a critical analysis of the relative strengths and limitations of each ITP implementation.

\subsubsection{Quantitative Analysis} \label{sec:quantitative-analysis}

The quantitative analysis focuses on three key metrics across different proof results:

\begin{itemize}
    \item \textbf{Lines of code.} The total number of lines of code (excluding code unrelated to proofs) in each file.
    \item \textbf{Number of auxiliary lemmas.} The number of helper lemmas required to support the proofs in each module.
    \item \textbf{External library dependencies.} The number and nature of external libraries relied upon by each module.
\end{itemize}

\autoref{tab:imandra-formal-results} gives an overview of the Imandra proof checker formalisation. The total codebase spans 2113 L.O.C of code and relies on 267 auxiliary lemmas to support the proof checker soundness. A significant portion of the proof effort is concentrated in two results: \texttt{ReLU splits are covering} and \texttt{Sound application of DNN polynomial Farkas lemma}, which together account for 1563 L.O.C., or 75\%~ of the total.

\autoref{tab:rocq-formal-results} summarises the Rocq formalisation. The implementation comprises 978 L.O.C supported by 52 auxiliary lemmas, organised across seven theorem modules. The most proof-intensive results, \texttt{Sound application of DNN polynomial Farkas lemma} and \texttt{ReLU splits are covering}, contribute 124 and 145 L.O.C. respectively. The proof structure relies heavily on MathComp's algebraic hierarchy for ring, matrix, and vector-space theories, which provides the foundational reasoning used throughout the formalisation.

The disparity in code size and lemma count between the two implementations stems primarily from differences in the underlying proof environments. Imandra's proof style encourages explicit, low-level reasoning about algebraic structures, requiring manually constructed induction schemes and detailed case analyses. Rocq, by contrast, leverages the MathComp library's algebraic hierarchy, which provides pre-verified theories for matrices and vector spaces.

This allows the Rocq formalisation to express algebraic constraints at a higher level of abstraction. For example, operations such as matrix multiplication and dot products are built into MathComp's matrix theory, whereas Imandra requires explicit recursive definitions and lemmas over real-number lists. This difference is reflected in the reduction from 2113 to 978 L.O.C. (approximately 54\%) and from 267 to 52 auxiliary lemmas (approximately 81\%), with the most substantial savings appearing in the algebraically-intensive modules.

Another reason that contributes to the considerable difference in terms of L.O.C. is the exploitation of dependent types in the Rocq implementation. Since Imandra does not have such a feature, it has to enforce the validity of its vectors recursively. An example of this is the \texttt{check\_tree\_soundness} theorem, as can be seen in the Imandra version in \autoref{code:imandra-check-tree-soundness} and the corresponding Rocq version in \autoref{code:check-tree-soundness}. The main difference between both is the usage of \texttt{well\_formed\_vector} and \texttt{well\_formed\_tableau\_bounds}, which ensure structural guarantees of the tableau and bounds size. These functions are not necessary in the Rocq code, since the guarantees are encoded in each type.

\begin{lstlisting}[language=Imandra, caption={Imandra theorem for \texttt{check\_tree\_soundness} with explicit well-formedness preconditions.}, label={code:imandra-check-tree-soundness}, numbers=left, breaklines=true]
theorem check_tree_soundness (tableau: real list list) (upper_bounds: real list) (lower_bounds: real list)
        (constraints: Constraint.t list) (tree: Proof_tree.t) (x: real list) =
    well_formed_vector tableau x
    && well_formed_tableau_bounds tableau upper_bounds lower_bounds
    && check_all_splits tree constraints
    && check_tree (mk_eq_constraints tableau) upper_bounds lower_bounds constraints tree
    ==>
    unsat tableau upper_bounds lower_bounds constraints x
[@@by [%expand "valid_proof"]
   @> [%expand "well_formed_vector"]
   @> induction ~id:[%id check_tree] ()
      @>>| [%use check_tree_soundness_full tableau upper_bounds lower_bounds constraints tree x]
             @> [%use check_tree_parent_imply_check_tree_children (mk_eq_constraints tableau)
                      upper_bounds lower_bounds constraints tree]
             @> [%use well_formed_preservation tableau upper_bounds lower_bounds (split_of_node tree)]
             @> auto
    ]
\end{lstlisting}

Interestingly, due to its high degree of automation, Imandra's main strength is not reducing total L.O.C. but rather reducing the amount of human-written proof code per theorem. Therefore, an Imandra user will likely spend more time specifying lemmas to guide the prover's automated reasoning than writing explicit proof scripts. This can be seen by comparing the L.O.C. to the number of auxiliary lemmas: Imandra's ratio is 7.9 L.O.C. per lemma (2113 / 267), whereas Rocq's ratio is 18.8 L.O.C. per lemma (978 / 52). Each Rocq lemma therefore contains more human-written proof code on average, because Rocq requires explicit tactic-level proofs while Imandra handles more of the low-level reasoning automatically through computation and induction.

The Rocq formalisation incurs a heavier dependency footprint: it relies on MathComp's comprehensive algebraic library, which introduces additional compilation and verification overhead. However, this dependency is well-established and widely used in the formal verification community, mitigating long-term maintainability risks. The Imandra formalisation, with its self-contained proofs and minimal external dependencies, may be easier to maintain in isolation, but at the cost of significantly more proof code.

\subsubsection{Qualitative Analysis}

The qualitative analysis compares the proof procedures of both implementations, focusing on the structural differences in how lemmas are composed.

Since both proof checkers share the same core structure, the analysis concentrates on one of the most reasoning-intensive modules: the leaf soundness and the underlying application of the DNN polynomial Farkas lemma. The leaf soundness module is particularly instructive because it highlights the main differences and trade-offs between both ITPs.

As shown in \autoref{code:leaf-soundness}, both lemmas prove that \texttt{check\_tree} implies \texttt{unsat} for leaf nodes. The main structural difference is that, as discussed in \autoref{sec:quantitative-analysis}, Imandra's statement includes explicit dimension checks (lines 5 and 6 in the Imandra code), whereas Rocq encodes these constraints in the matrix types (lines 2 and 3 in the Rocq code) and therefore omits them from the lemma statement.

\begin{figure}[h]
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Coq, breaklines=true, numbers=left, basicstyle=\tiny\ttfamily]
Lemma leaf_soundness
  (tableau : (m.+2).-tuple ('rV[R]_n))
  (ub lb : 'rV[R]_n)
  (constraints : seq Constraint.t)
  (contradiction : (m.+2).-tuple R)
  (x : 'rV[R]_n) :
  let system' := Cert.mk_eq_constraints tableau in
  Checker.check_tree system' ub lb constraints (ProofTree.leaf contradiction) ->
  Sat.unsat tableau ub lb constraints x.
Proof.
  intros system' H.
  move: (not_eval_system_implies_unsat tableau ub lb constraints contradiction x) => H_eval.
  move: (check_cert_implies_not_eval_system system' ub lb constraints contradiction x) => H_test.
  auto.
Qed.
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Imandra, numbers=left, breaklines=true, basicstyle=\tiny\ttfamily]
lemma soundness_leaf tableau upper_bounds lower_bounds relu_constraints x proof_tree =
    match proof_tree with
    | Node _ -> true
    | Leaf contradiction ->
        well_formed_tableau_bounds tableau upper_bounds lower_bounds &&
        List.length x = List.length (List.hd tableau) &&
        check_tree (mk_eq_constraints tableau) upper_bounds lower_bounds relu_constraints proof_tree
        ==> 
        unsat tableau upper_bounds lower_bounds relu_constraints x
[@@by [%use check_tree_implies_check_cert (mk_eq_constraints tableau) upper_bounds lower_bounds relu_constraints proof_tree]
   @> [%use not_eval_system_implies_unsat tableau upper_bounds lower_bounds relu_constraints x]
   @> auto]
\end{lstlisting}
\end{minipage}
\caption{Comparison of \texttt{leaf\_soundness} in Rocq (left) and Imandra (right).}
\label{code:leaf-soundness}
\end{figure}

The deeper difference lies in how the intermediate lemmas are proved.  Specifically, how the \texttt{not\_eval\_system\_implies\_unsat} lemma, which bridges the certificate-checking result (the contradiction system evaluates to false) into the formal \texttt{unsat} predicate used by the rest of the proof tree soundness theorem, is proven. The lemma is split into two independent directions: equality and inequality directions.

\autoref{code:system-soundness-eq} compares the equality direction.  The Imandra lemma \texttt{tableau\_reduction\_\\soundness} is proved by induction on the tableau (a list of lists).  The Rocq lemma \texttt{system\_soundness\_eq} uses \texttt{all\_tnthP} to iterate over the typed tuple of rows and \texttt{dot\_product\_eq\_kernel} (itself proved via MathComp's \texttt{mul\_row\_col}) to rewrite each row evaluation into the kernel condition. No explicit induction is needed because the type \texttt{m.-tuple ('rV[R]\_n)} guarantees the shape.

\begin{figure}[h]
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Coq, numbers=left, breaklines=true, basicstyle=\tiny\ttfamily]
Lemma dot_product_eq_kernel
    (hd : 'rV[R]_n) (x : 'rV[R]_n) :
  Arithmetic.dot_product
    (Arithmetic.rv_addn1_succ
      (row_mx hd 0))
    (Arithmetic.cv_addn1_succ
      (col_mx (trmx x) 1%:M))
  = Arithmetic.dot_product
      hd (trmx x).
Proof.
  rewrite /Arithmetic.dot_product
    /Arithmetic.rv_addn1_succ
    /Arithmetic.cv_addn1_succ
    !castmx_id
    (mul_row_col hd 0 x^T 1%:M)
    mul0mx addr0.
Qed.

Lemma system_soundness_eq
  (tableau : (m.+2).-tuple
    ('rV[R]_n)) (x : 'rV[R]_n) :
  Arithmetic.is_in_kernel tableau x ->
  let sys := Cert.mk_eq_constraints
    tableau in
  FarkasSoundness.eval_system
    sys (trmx x).
Proof.
  move=> Hin. apply/all_tnthP => i.
  rewrite tnth_map
    /FarkasSoundness.eval_expr.
  have Hi := (all_tnthP Hin) i.
  by rewrite dot_product_eq_kernel.
Qed.
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Imandra, numbers=left, breaklines=true, basicstyle=\tiny\ttfamily]
lemma tableau_reduction_soundness tableau x =
    List.length x =
      List.length (List.hd tableau) &&
    well_formed_tableau tableau &&
    is_in_kernel tableau x
    ==>
    eval_system (mk_eq_constraints
      tableau) x
[@@by induct ~on_vars:["tableau"] ()]
\end{lstlisting}
\end{minipage}
\caption{Comparison of the equality direction of the system soundness proof between Rocq (left) and Imandra (right).}
\label{code:system-soundness-eq}
\end{figure}

\autoref{code:system-soundness-geq} compares the inequality direction.  The Imandra lemma \texttt{mk\_geq\_constraints\\\_reduction\_soundness} chains multiple auxiliary lemmas, each proved by induction on the bounds list.  The Rocq lemma \texttt{system\_soundness\_geq} uses \texttt{forallP} to unpack the predicate, splits the goal with \texttt{all\_cat}, and discharges each half by iterating over indices with \texttt{all\_tnthP} and applying \texttt{subr\_ge0}. 

\begin{figure}[h]
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Coq, numbers=left, breaklines=true, basicstyle=\tiny\ttfamily]
Lemma system_soundness_geq
  (ub lb : 'rV[R]_n) (x : 'rV[R]_n) :
  Arithmetic.bounded x ub lb ->
  FarkasSoundness.eval_system
    (Cert.mk_geq_constraints ub lb)
    (trmx x).
Proof.
  move=> /forallP Hb.
  rewrite all_cat; apply/andP; split.
  - refine (introT
      (all_tnthP _) _) => i.
    rewrite tnth_mktuple
      /FarkasSoundness.eval_expr
      dot_product_mk_bound_poly.
    have /andP[_ Hu] := Hb i.
    by rewrite mulN1r addrC subr_ge0.
  - refine (introT
      (all_tnthP _) _) => i.
    rewrite tnth_mktuple
      /FarkasSoundness.eval_expr
      dot_product_mk_bound_poly.
    have /andP[Hl _] := Hb i.
    by rewrite mul1r subr_ge0.
Qed.
\end{lstlisting}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\begin{lstlisting}[language=Imandra, numbers=left, breaklines=true, basicstyle=\tiny\ttfamily]
lemma mk_geq_constraints_reduction_soundness
    ubs lbs xs =
    List.length xs = List.length ubs &&
    List.length xs = List.length lbs &&
    bounded xs ubs lbs
    ==>
    eval_system
      (mk_geq_constraints ubs lbs) xs
[@@by [%use bounded_upper xs ubs lbs]
   @> [%use bounded_lower xs ubs lbs]
   @> [%use ...eval ubs xs]
   @> [%use ...eval lbs xs]
   @> [%use eval_system_concat_invariant]
   @> auto]
\end{lstlisting}
\end{minipage}
\caption{Comparison of the inequality direction of the system soundness proof between Rocq (left) and Imandra (right).}
\label{code:system-soundness-geq}
\end{figure}

Even though Imandra proofs seem much simpler than Rocq's due to the brevity of their proof steps, the reality is that both lemmas rely on multiple auxiliary lemmas that might not be explicitly referenced but still need to be defined. This fact contributes to the minimal length of the proofs, since the reasoning is performed by Imandra's automated reasoning system. However, as seen in \autoref{sec:quantitative-analysis}, the codebase tends to bloat due to the excessive number of auxiliary lemmas that need to be stated. In contrast, Rocq proofs are longer and more complex but depend on fewer auxiliary lemmas, making the reasoning underlying their soundness more explicit and easier to understand.



