\section{Implementation} \label{sec:implementation}

The implementation of the certified proof checker consists of three main components:
\begin{itemize}
    \item \textbf{Certificate Parser.} A Haskell program that reads proof certificates in JSON format and generates Rocq source files containing equivalent Gallina definitions.
    \item \textbf{Proof Checker.} A Rocq program that verifies whether a parsed proof certificate constitutes a valid proof of unsatisfiability.
    \item \textbf{Proof Checker Soundness.} A formal proof in Rocq establishing that if the proof checker accepts a certificate, then the original verification query is genuinely unsatisfiable.
\end{itemize}

\autoref{fig:system-workflow} provides an overview of the workflow of the developed system. When Marabou determines that a verification query is UNSAT, it produces a proof certificate of the result in JSON format. This certificate is passed to the certificate parser, which converts it into Rocq source files containing equivalent Gallina definitions. The generated files are then imported by the proof checker where it verifies the certificate. Finally, the soundness theorem establishes that any certificate accepted by the proof checker constitutes a valid UNSAT proof of the corresponding verification result.

\input{components/diagrams/system_architecture.tex}

\subsection{Certificate Parser}
To reason about the proof certificates produced by Marabou in the UNSAT case, they must first be imported into the proof checker. In the original Imandra development, this is achieved within the same program: since Imandra is embedded in a subset of OCaml, it can leverage OCaml parsing libraries to read and translate the proof certificates directly at execution time. Consequently, no external preprocessing stage is required.

This approach is not directly transferable to Rocq. The proof checker developed in this work is implemented in Gallina, which cannot invoke arbitrary OCaml parsing code during proof checking. Therefore, a separate preprocessing stage was introduced. A Haskell program parses the JSON proof certificates produced by Marabou and generates Rocq source files containing equivalent Gallina definitions. These generated files are then imported into the proof checker, allowing it to reason about the certificates using native Rocq data structures.

Marabou's proof certificates are not organised in a way that is convenient for recursive verification. In particular, the tightening information is associated with parent nodes, whereas the proof checker benefits from a representation in which this information is readily available when processing each proof step. Consequently, before verification, the parsed certificates should be translated into data types better suited to the recursive structure of the proof checker.

\todo[inline]{Should I be mentioning the information below here or should I do it in the "Evaluation" section?} The point at which this translation is performed differs between the two implementations. In the original Imandra development, the transformation is carried out within the Imandra code after the certificates have been parsed. In this work, the transformation is instead performed by the Haskell parser while generating the Rocq files. As a result, the Rocq implementation operates directly on the representation expected by the proof checker, keeping the Gallina code focused solely on certificate verification.


\subsection{Proof Checker}


The proof checker has exactly one job: to determine whether an UNSAT proof certificate genuinely certifies unsatisfiability. To do so, it operates over the following data, all of which is obtained from the certificate parser in a form of Gallina files. 

\begin{itemize}
    \item \textbf{Upper and lower bounds.} Vectors $u$ and $l$ assigning an admissible interval to each variable. These define the search space of the verification query.
    \item \textbf{Constraints.} A collection of linear constraints representing the network's weights, biases, and the property under verification, alongside the ReLU constraints arising from activation functions.
    \item \textbf{Tableau.} A matrix $T$ encoding the linear relationships between basic and non-basic variables, as maintained by Marabou's simplex engine during the search.
    \item \textbf{Proof tree.} A binary tree whose internal nodes record case splits (typically over ReLU phases) and whose leaves carry Farkas coefficients that witness the unsatisfiability of the leaf's sub-problem.
\end{itemize}

As mentioned earlier, the Marabou DNN verifier produces proof certificates in a tree-like structure that the proof checker traverses recursively. \autoref{code:proof-tree} shows the formalisation of Marabou proof trees as an inductive data type. It consists of two constructors: a \texttt{node} and a \texttt{leaf}. Each leaf holds the Farkas contradiction vector, while each node holds the split information required to tighten the bounds, together with references to the left and right child subtrees.

\begin{lstlisting}[language=Coq, caption={Marabou proof tree as an inductive data type in Rocq.}, label={code:proof-tree}]
Inductive t : Type :=
  (*  split information, left child, right child *)
  | node : Split.t -> t -> t -> t
  (* contradiction vector *)
  | leaf : (m.+2).-tuple R -> t.
\end{lstlisting}

The \texttt{check\_tree} function, shown in \autoref{code:check-tree}, recursively traverses the populated proof tree. At each leaf, the proof checker verifies that the provided Farkas coefficients constitute a valid witness of unsatisfiability for the linear constraints under the current bounds. At each internal node, it first tightens the bounds and then checks that the current split is consistent with the constraints before recursively verifying both child subtrees.

\begin{lstlisting}[language=Coq, caption={Proof checker tree verifier.}, label={code:check-tree}]
Fixpoint check_tree
  (tableau : system m n)
  (ub lb : 'rV[R]_n)
  (constraints : seq Constraint.t)
  (proof_node : ProofTree.t)
  : bool :=
  match proof_node with
  | ProofTree.leaf contradiction =>
    check_contradiction contradiction tableau ub lb
  | ProofTree.node split tleft tright =>
    let bounds := Split.update_bounds_from_split ub lb split in
    let valid_split := Split.check_split split constraints in
    let valid_left  := check_tree tableau bounds.1.2 bounds.1.1 constraints tleft in
    let valid_right := check_tree tableau bounds.2.2 bounds.2.1 constraints tright in
    valid_split && valid_left && valid_right
  end.
\end{lstlisting}

TODO: DIVE DEEPER INTO THE PROOF CHECKER SYSTEM.
% The proof checker makes use of dependent types all over the the program.

\subsection{Proof Checker Soundness}

TODO

